---
title: "Homework 3 193DS"
author: "Valentina Caro"
date: "03-05-2026"
format: html
editor: visual
execute:
  message: false 
  warning: false
---

## Set Up

```{r}
#| label: reading in packages + data

library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(scales)
library(ggeffects)
library(MuMIn)

salinity <- read_csv( # creating data frame
  here(
    "data",
    "salinity-pickleweed.csv"
  ))

```

# Problem 1.

## a) Appropriate Test

In order to determine the strength of the relationship between salinity (mS/cm) and California pickleweed biomass (g), I would use an Ordinary Least Squares (OLS) test and a Pearson's r test. An OLS provides a model equation that describes the linear relationship between a response variable and predictor variable, while a Pearson's r test tell us the strength and directionality of this relationship.

## b) Visualization

```{r}
#| label: scatter plot

ggplot(data = salinity,# setting up ggplot using data frame "salinity"
       aes(x = salinity_mS_cm,# setting x axis
           y = pickleweed)) + # setting y axis
  geom_point(color = "darkorchid4") + # changing point color
  labs(x = "Salinity (mS/cm", # labelling x axis
       y = "Pickleweed Biomass (g)", # labelling y axis
       title = "As Salinity Increases, Biomass Increases") + # making a title
  theme(plot.title.position = "plot") + # changing title position
  theme_minimal() # changing theme to minimal

```

## c) Checking Assumptions and Test

```{r}
#| label: model fitting

salinity_lm <- lm(
  pickleweed ~ salinity_mS_cm,
  data = salinity
) # created linear regression model using salinity data frame

summary(salinity_lm) # display model

```

```{r}

salinity_pred <- ggeffects::ggpredict(salinity_lm, # making predictions based on linear model
                           terms = "salinity_mS_cm") 
 
```

```{r}
#| label: residual plots

par(mfrow = c(2,2)) # making residual plots to check assumptions
plot(salinity_lm)

```

Checking Assumptions:

The diagnostic plots look good. Residuals look homoscedastic (constant variance) based on residuals vs fitted plot and scale location plot. Residuals (errors) appear to be normally distributed based on q-q plot.

```{r}
#| label: pearson's r

cor.test(salinity$salinity_mS_cm, salinity$pickleweed,
         method = "pearson")

```
