---
title: "Homework 3 193DS"
author: "Valentina Caro"
date: "03-05-2026"
format: html
editor: visual
execute:
  message: false 
  warning: false
---

## Set Up

```{r}
#| label: reading in packages + data

library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(scales)
library(ggeffects)
library(MuMIn)

salinity <- read_csv( # creating data frame
  here(
    "data",
    "salinity-pickleweed.csv"
  ))

```

# Problem 1.

## a) Appropriate Test

In order to determine the strength of the relationship between salinity (mS/cm) and California pickleweed biomass (g), I would use an Ordinary Least Squares (OLS) test and a Pearson's r test. An OLS provides a model equation that describes the linear relationship between a response variable and predictor variable, while a Pearson's r test tell us the strength and directionality of this relationship.

## b) Visualization

```{r}
#| label: scatter plot

ggplot(data = salinity,# setting up ggplot using data frame "salinity"
       aes(x = salinity_mS_cm,# setting x axis
           y = pickleweed)) + # setting y axis
  geom_point(color = "darkorchid4") + # changing point color
  labs(x = "Salinity (mS/cm", # labelling x axis
       y = "Pickleweed Biomass (g)", # labelling y axis
       title = "As Salinity Increases, Biomass Increases") + # making a title
  geom_smooth(method = "lm", # adding linear model line
              color = "magenta") +
  theme(plot.title.position = "plot") + # changing title position
  theme_minimal() # changing theme to minimal

```

## c) Checking Assumptions and Test

```{r}
#| label: least squares model

salinity_lm <- lm(
  pickleweed ~ salinity_mS_cm,
  data = salinity
) # created linear regression model using salinity data frame

summary(salinity_lm) # display model

```

```{r}
#| label: residual plots

par(mfrow = c(2,2)) # making residual plots to check assumptions
plot(salinity_lm)

```

Checking Assumptions:

The diagnostic plots look good. Residuals look homoscedastic (constant variance) based on residuals vs fitted plot and scale location plot. Residuals (errors) appear to be normally distributed based on q-q plot. Linearity can be assumed based on scatterplot including line (seen above).

## d) Results

To evaluate the strength in the relationship between pickleweed biomass and soil salinity, I chose to use a least squares regression model. Using this model, I found evidence to suggest that soil salinity affected pickleweed biomass. (ordinary least squares regression: F(1,21) = 8.4, p = 0.009, $\alpha$ = 0.05, adjusted R\^2 = 0.25)

With each 1 mS/cm increase in soil salinity, pickleweed biomass increased by 2.01 +/- 0.72 g. This model explains 25% of the variatation in pickleweed biomass.

## e) Implications

The results of my linear model provide evidence that as soil salinity increases, pickleweed biomass increases. Planting pickleweed in areas with higher soil salinity may increase the plant's success, however, there are some limitations to this study. First, there is likely a maximum threshold for soil salinity that if surpassed, may decrease pickleweed plant growth, so I would not recommend planting pickleweed in areas with a higher soil salinity than 10 mS/cm. Additionally, an R\^2 of 0.25 is low, suggesting there are many other variables that affect pickleweed biomass and require further testing to determine the most ideal sites for restoration.

## 
