---
title: "Homework 3 193DS"
author: "Valentina Caro"
date: "03-05-2026"
format: html
editor: visual
execute:
  message: false 
  warning: false
---

## Set Up

```{r}
#| label: reading in packages + data

library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(scales)
library(ggeffects)
library(MuMIn)

salinity <- read_csv( # creating data frame
  here(
    "data",
    "salinity-pickleweed.csv"
  ))

```

# Problem 1.

a\) Appropriate Test

In order to determine the strength of the relationship between salinity (mS/cm) and California pickleweed biomass (g), I would use an Ordinary Least Squares (OLS) test and a Pearson's r test. An OLS provides a model equation that describes the linear relationship between a response variable and predictor variable, while a Pearson's r test tell us the strength and directionality of this relationship.

b\) Visualization

```{r}
#| label: visualization

ggplot(data = salinity,# setting up ggplot using data frame "salinity"
       aes(x = salinity_mS_cm,# setting x axis
           y = pickleweed)) + # setting y axis
  geom_point(color = "darkorchid4") + # changing point color
  labs(x = "Salinity (mS/cm", # labelling x axis
       y = "Pickleweed Biomass (g)", # labelling y axis
       title = "As Salinity Increases, Biomass Increases") + # making a title
  theme(plot.title.position = "plot") + # changing title position
  theme_minimal() # changing theme to minimal

```
